---
title: "Writen Assignment Answers"
author: "Adam Armeier"
date: "6/26/2022"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(sf)
library(spdep)
library(rvest)
library(knitr)

earthquakes <- read_xlsx("Data\\Data Set - Earthquakes.xlsx")[-1, ] %>% 
  dplyr::filter(is.na(Longitude) == FALSE) %>% 
  st_as_sf(coords = c('Longitude', 'Latitude'), remove = F, crs = 4326)

countries <- st_read("Data\\ne_110m_admin_0_countries\\ne_110m_admin_0_countries.shp", crs = 4326)

# Output file from embassy_scrape.R
embassies <- read_csv("Data\\embassies_consulates.csv")

#fault_lines <- st_read("https://raw.githubusercontent.com/GEMScienceTools/gem-global-active-faults/master/geojson/gem_active_faults_harmonized.geojson")

# Read in Natural Earth populate places file to get coordinates of cities
# Many manual naming changes to ensure the webscraped embassies match the
# spelling in the Natural Earth file
pop_places <- st_read("Data\\ne_10m_populated_places_simple\\ne_10m_populated_places_simple.shp", crs=4326) %>%
  dplyr::mutate(nameascii = as.character(nameascii)) %>% 
  dplyr::filter(sov0name != "United States") %>% 
  dplyr::mutate(nameascii = 
                  case_when(
                    nameascii == "Kobenhavn" ~ "Copenhagen",
                    nameascii == "Port-of-Spain" ~ "Port of Spain",
                    nameascii == "Tel Aviv-Yafo" ~ "Tel Aviv",
                    nameascii == "Basra" ~ "Basrah",
                    nameascii == "Willemstad" ~ "Curacao",
                    nameascii == "Az Zahran" ~ "Dhahran",
                    nameascii == "Dusseldorf" ~ "Düsseldorf",
                    TRUE ~ nameascii
                  )) %>% 
  dplyr::filter(featurecla != "Populated place" | 
                  nameascii == "Ciudad Juarez" | 
                  nameascii == "Montreal" |
                  nameascii == "Calgary" |
                  nameascii == "Dhahran" | 
                  nameascii == "Erbil" |
                  nameascii == "Frankfurt") %>% 
  dplyr::filter(!(nameascii == "La Paz" & 
                    (sov0name == "Mexico" | sov0name == "Honduras")))%>% 
  dplyr::filter(!(nameascii == "Santiago" & 
                    (sov0name == "Panama" | 
                       sov0name == "Dominican Republic")))

# Join the populated places and embassies file
sf_df <- left_join(pop_places, 
                   embassies, 
                   by = c("nameascii" = "Location")) %>% 
  dplyr::filter(!is.na(Type)) %>% 
  arrange(nameascii)

# Created a global hex-grid with unique indexes in each hexagon
sfc_global <- st_bbox(pop_places) %>% st_as_sfc()
global_grid <- st_make_grid(sfc_global, 2, crs = st_crs(sf_df), square = F)
global_grid <- st_sf(hexid = 1:length(lengths(global_grid)), global_grid)

# Get the intersection between the hexagons and earthquake events
hexid <- sapply(st_intersects(x = earthquakes, y = global_grid), 
                function(z) if (length(z)==0) NA_integer_ else z[1])
earthquakes$hexid <- hexid

# Provide summary statistics for each hexagon
earthquake_hexsum <- earthquakes %>% group_by(hexid) %>% 
  summarize(events = n(),
            events_maj = sum(Mag >= 7),
            Deaths = sum(Deaths, na.rm = T),
            Injuries = sum(Injuries, na.rm = T)) %>% 
  st_set_geometry(NULL)

# Join the initial grid file and the summarized statistics
quake_grid <- left_join(global_grid, earthquake_hexsum) %>% 
  replace_na(list(events = 0, Deaths = 0, Injuries = 0))

# Calculate risk based on the number of events and deaths in an area
quake_grid <- quake_grid %>% 
  mutate(risk = log(events + 1) * log(Deaths + 1),
         risk_lvl = ifelse(risk < 5, 1,
                           ifelse(risk < 15, 2, 3))) %>% 
  dplyr::filter(risk > 0)

# More complicated analysis based upon regional effects


# Get the intersection between the quake grid and the embassies file
sf_df$hexid <- sapply(st_intersects(x = sf_df, y = global_grid),
                        function(z) if (length(z)==0) NA_integer_ else z[1])
quake_grid_simple <- quake_grid
st_geometry(quake_grid_simple) <- NULL
embassy_risk <- left_join(sf_df, quake_grid_simple) %>% 
  replace_na(list(risk = 0, risk_lvl = 0))
```

## Question 1 - Study Design

Please discuss in detail your approach to this analysis.  Study Design may Include how you *defined the question*, data, methodology, and analytic design or your methodology to data exploration and explanation, the assumptions taken, how you would ensure the integrity and accuracy of any results, and any considerations for communication to a varied audience. 

I defined the question directly from the provided scenario, which asks for 
a "rapid preliminary analysis of the risk of earthquakes to our overseas
missions." Therefore, I consider the research questions to be 
straightforward from the scenario: "What is the risk of earthquakes to 
State Department overseas missions?" This requires defining a method to
measure risk of earthquakes and comparing it for each of the State 
Department's overseas missions. 

I choose to use a method that defines the relative, rather than the absolute
risk for the State Department's missions. My assumption is that this 
preliminary assessment is to provide decision makers with information on  
which missions are most at risk, rather than provide a likelihood that a 
particular mission will be affected by a serious earthquake. To provide an
absolute risk, I would probably develop a logistical regression 
model to attempt to quantify the risk based upon a number of factors 
related to earthquake risk. However, since I know very little about 
earthquake risk and know that earthquake prediction has not been very 
succesful, I rely on a straightforward assessment based on the provided 
earthquake spreadsheet.

Although not feasible for this scenario, I would also try to clarify  
how the person or organizaiton making the request defines risk. I would 
likely try to communicate with the customer to ensure that I understand 
their desired output. For example, do they perceive the risk of earthquakes 
to simply be a risk of an earthquake event occuring, or does it also need 
to account for factors such as the capabilities of a host nation to provide 
disaster relief support, the ability of the Department of State or other US 
government support entities to access the mission in an emergency, or the 
risk of factors like tidal waves that may result from an earthquakes. I 
also looked at the Joint Strategic Plan for guidance on the type of risk 
assessement mentioned. Given the text from the prompt and Performance Goal 
4.3.4 to "protect personnel through advanced building" I assume the goal is
to prioritize which missions are most at risk and may need to be rebuilt or
reinforced in order to protect the personnel stationed there. Hence, I use
a relative risk assessment.

I use a spatial approach of analysis to account for variations in the 
threat level to consulates within countries where the State Department may 
have multiple consulates. For example, analyzing data for the United States 
would provide a much different earthquake risk when comparing, for example, 
California to Illinois. The spatial methodology allows for accounting for 
the variation within various countries, particularly those with consulates 
outside of the capital that may have varying risks to eathquakes. The 
features in the dataset that are the most complete are the time, the 
geographic coordinates, the magnitude, and the focal depth. Otherwise, most
of the spreadsheet contains missing data for about half of the features.

I decide to compile a geographic index that accounts for the frequency of 
events and the number of fatalities in the events. One flaw of using the
provided dataset is that it doesn't account well for low likelihood, high
impact events. An area

Part of the reason I use a simple index is to ensure that the results are
easily explicable to a senior policymaker, particularly if they are 
presented as one simple graphic. Since the prompt suggests the Secretary is
requesting assessments about multiple natural disasters, I assume that there
would be limited time to discuss the results and the methods that went into
the assessment. I presume the assment would likely be discussed with 
assessments of hurricane risk, flooding risk, and other natural disaster 
risks.


## Question 2 - Data Analysis

*What are key findings found in your analysis? Which embassies are most at risk? Are there groups of embassies that share similar features or profiles with respect to earthquake risk? Were there outliers or areas with data quality issues? Please describe how you conducted your data analysis, including which tools you used, why you chose them, and data preprocessing steps you took, and justify the results of your analysis.*   

I find that the embassies and consulates most at risk of being affected by 
an earthquake are along or near the Pacific Coast in Central America and
east Asia, in southeastern Europe. Port-au-Prince is the most at risk, 
primarily because of the effect of the 2010 earth quake. Tbilisi and Kabul, 
are in mountainous areas that experience frequent earthquakes. 

```{r}
subset_emb <- embassy_risk %>% slice_max(risk, n = 10) %>% 
  arrange(desc(risk)) %>% 
  select(name, sov0name, risk, risk_lvl) %>% 
  rename(City = name, Country = sov0name,
         Risk = risk, `Risk Level` = risk_lvl )
st_geometry(subset_emb) <- NULL

kable(subset_emb)
```

I consider the primary data quality issue to be the simplicity of the 
analysis, the geographic scale for the results, and data that is missing for
before 1970. I decided to utilize a relatively simple model because there 
was so much categorical data missing within many of the events. Only the 
coordinates, magnitude, and focal depth were available for almost all of 
the events. I chose to use the categories for *Deaths* to *Injuries* 
to measure the human effect of the various events, despite there being 
many missing values. A quick visualization of the location of events with
no recorded deaths or injuries showed that they were often over water or
sparsely populated areas. This gives some assurance that missing data 
actually reflects a 0 value, rather than simply being missing data.

With more time, I would have liked to define more precise characteristics 
related to geographic scale. I used a small-scale analysis because of the 
global scale of analysis and because of the low-precision of the Embassy 
data that I had. Without a geographic file of embassy locations, I attempted
various methods to generate a proper file. I tried to acquire the data from
OpenStreetMap, but it was very imprecise. I decided to scrape a State 
Department website that listed all the cities globally with an Embassy or
Consolate and then I joined that file to a geojson file that lists major
global cities. This method was imprecise and required some manual 
adjustments. In addition, it means that the analysis is done at the city
level, rather than at the precise embassy geographic coordinates.

Although a rather minor point, the lack of data before 1970 means that the
model doesn't account for low likelihood, high impact events in areas where
fault lines may be overdue for an event. With more time, I would like to 
coolaborate with academics or colleagues at USGS to discuss how important 
this factor would be.

I chose to do my analysis in R, primarily using the tidyverse and sf 
packages. Although I could have done this analysis in ArcGIS, I wanted to
use R in order to accomodate better reproducability if other analysts 
wanted to leverage this model. Given that this analysis intended to be 
preliminary, it is likely to be modified and refined with more time. With
a desire to work with academic researchers or outside consultants, I find 
that using GitHub would provide a good way to share the analysis and data
with other potential collaborators. For example, I have uploaded my files 
to GitHub if you would like to see the example of the code that I used for
this document
(https://github.com/Slushmier/earthquake-data).

I have also come to prefer using R for spatial analysis over ArcGIS as it
requires more attention to the spatial parameters that are required for an
analysis.


## Question 3 - Communication

*Please describe in detail how you would communicate your results to a diverse, non-technical audience. Provide an executive summary of your analysis for senior management which does not exceed 500 words.* 

Wherever plausible, I prefer to use a map to display results because of the
ability to 

```{r}
ggplot() + 
  geom_sf(data = countries) +
  geom_sf(data = quake_grid, aes(fill = risk_lvl), alpha = 0.2) +
  geom_sf(data = embassy_risk, aes(color = risk_lvl))

### Check that zero death & injury events are remote
# earthquakes_sub <- earthquakes %>% 
#  dplyr::filter(is.na(Deaths) & is.na(Injuries))
# ggplot() + geom_sf(data = countries) + 
#  geom_sf(data = earthquakes_sub, aes(color = Mag))
```


## Question 4 - Validation

*Hypothetically, if a member of your team completed this work using software or methodology in which you are not proficient, describe how you would validate their work.* 

Validation can be difficult when you are unclear about the tool or 
methodology that may be used by another analyst. I choose to put my analysis
on GitHub (https://github.com/Slushmier/earthquake-data) so that anyone 
reviewing the 

I could also compare the results to their alignment with similar academic
models. For example, I validated my results against the Global Seismic 
Hazard Assessment program's results.